---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidyposterior

<!-- badges: start -->
[![R build status](https://github.com/tidymodels/tidyposterior/workflows/R-CMD-check/badge.svg)](https://github.com/tidymodels/tidyposterior/actions)
[![Codecov test coverage](https://codecov.io/gh/tidymodels/tidyposterior/branch/master/graph/badge.svg)](https://codecov.io/gh/tidymodels/tidyposterior?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/tidyposterior)](https://CRAN.r-project.org/package=tidyposterior)
[![Downloads](http://cranlogs.r-pkg.org/badges/tidyposterior)](https://CRAN.r-project.org/package=tidyposterior)
![](https://img.shields.io/badge/lifecycle-maturing-blue.svg)
<!-- badges: end -->

This package can be used to conduct _post hoc_ analyses of resampling results generated by models. 

For example, if two models are evaluated with the root mean squared error (RMSE) using 10-fold cross-validation, there are 10 paired statistics. These can be used to make comparisons between models without involving a test set. 

There is a rich literature on the analysis of model resampling results such as McLachlan's [_Discriminant Analysis and Statistical Pattern Recognition_](https://books.google.com/books?id=O_qHDLaWpDUC&lpg=PR7&ots=6GJnIREXZM&dq=%22Discriminant%20Analysis%20and%20Statistical%20Pattern%20Recognition%22&lr&pg=PR7#v=onepage&q=%22Discriminant%20Analysis%20and%20Statistical%20Pattern%20Recognition%22&f=false) and the references therein. This package follows _the spirit_ of [Benavoli _et al_ (2017)](https://people.idsia.ch//~marco/papers/2017jmlr-tests.pdf). 

tidyposterior uses Bayesian generalized linear models for this purpose and can be considered an upgraded version of the [`caret::resamples()`](https://topepo.github.io/caret/model-training-and-tuning.html#exploring-and-comparing-resampling-distributions) function. The package works with [rsample](https://rsample.tidymodels.org/) objects natively but any results in a data frame can be used. 

See [Chapter 11](https://www.tmwr.org/compare.html) of [_Tidy Models with R_](https://www.tmwr.org) for examples and more details. 

## Installation

You can install the released version of tidyposterior from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("tidyposterior")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tidymodels/tidyposterior")
```
## Example

To illustrate, here are some example objects using 10-fold cross-validation for a simple two-class problem: 

```{r setup, results = "hide"}
library(tidymodels)
library(tidyposterior)

data(two_class_dat, package = "modeldata")

set.seed(100)
folds <- vfold_cv(two_class_dat)
```

We can define two different models (for simplicity, with no tuning parameters).

```{r model-specs}
logistic_reg_glm_spec <-
  logistic_reg() %>%
  set_engine('glm')

mars_earth_spec <-
  mars(prod_degree = 1) %>%
  set_engine('earth') %>%
  set_mode('classification')
```

For tidymodels, the [tune::fit_resamples()] function can be used to estimate performance for each model/resample:

```{r tm-resamples}
rs_ctrl <- control_resamples(save_workflow = TRUE)

logistic_reg_glm_res <- 
  logistic_reg_glm_spec %>% 
  fit_resamples(Class ~ ., resamples = folds, control = rs_ctrl)

mars_earth_res <- 
  mars_earth_spec %>% 
  fit_resamples(Class ~ ., resamples = folds, control = rs_ctrl)
```

From these, there are several ways to pass the results to the `perf_mod()` function. The most general approach is to have a data frame with the resampling labels (i.e., one or more id columns) as well as columns for each model that you would like to compare. 

For the model results above, [tune::collect_metrics()] can be used along with some basic data manipulation steps: 

```{r df-results}
logistic_roc <- 
  collect_metrics(logistic_reg_glm_res, summarize = FALSE) %>% 
  dplyr::filter(.metric == "roc_auc") %>% 
  dplyr::select(id, logistic = .estimate)

mars_roc <- 
  collect_metrics(mars_earth_res, summarize = FALSE) %>% 
  dplyr::filter(.metric == "roc_auc") %>% 
  dplyr::select(id, mars = .estimate)

resamples_df <- full_join(logistic_roc, mars_roc, by = "id")
resamples_df
```

We can then give this directly to `perf_mod()`: 

```{r df-mod}
set.seed(101)
roc_model_via_df <- perf_mod(resamples_df, iter = 2000)
```

From this, the posterior distributions for each model can be obtained from the `tidy()` method: 

```{r post}
roc_model_via_df %>% 
  tidy() %>% 
  ggplot(aes(x = posterior)) + 
  geom_histogram(bins = 40, col = "blue", fill = "blue", alpha = .4) + 
  facet_wrap(~ model, ncol = 1) + 
  xlab("Area Under the ROC Curve")
```


See `contrast_models()` for how to analyze these distributions 


## Contributing

This project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

- For questions and discussions about tidymodels packages, modeling, and machine learning, please [post on RStudio Community](https://community.rstudio.com/new-topic?category_id=15&tags=tidymodels,question).

- If you think you have encountered a bug, please [submit an issue](https://github.com/tidymodels/tidyposterior/issues).

- Either way, learn how to create and share a [reprex](https://reprex.tidyverse.org/articles/articles/learn-reprex.html) (a minimal, reproducible example), to clearly communicate about your code.

- Check out further details on [contributing guidelines for tidymodels packages](https://www.tidymodels.org/contribute/) and [how to get help](https://www.tidymodels.org/help/).

